#machine learning
#project 1



#####Question: Can a prediction model be built using the training data the successfully predicts predict the manner in which HAR test subjests did their exercise?

#####Data: Data and a description of the data can be found at http://groupware.les.inf.puc-rio.br/har.  The training and test datasets must be downloaded and stored in working directory. The predictor is in the classe variable, A = proper implementation of the exercise while B to E are different improper methods.  


```{r, cache=TRUE}
train <- read.csv("./train.csv", na.strings=c("NA","#DIV/0!", ""))
test <- read.csv("./test.csv", na.strings=c("NA","#DIV/0!", ""))
dim(train); dim(test)
```


#####Features: Blanks, "#DIV/0!" were replaced with "NA" during data import.  Below, columns that contained no values where removed.  Columns 1 through 7 containing rowname, subject name, and date variables were eliminated since they will not contribute to predict.
Cross-validation was used by splitting the training data into training and test data set.  These sets will be used to evaluate models built with training on the test subset.  

```{r}
train2 <- train[,colSums(is.na(train)) == 0]
train3 <- train2[,-c(1:7)]

test2 <- test[,colSums(is.na(test)) == 0]
test3 <- test2[,-c(1:7)]
```



#####Algorithms: Rpart and randomForest were used to predict classe on subset of training data.


Install packages and load libraries.  Messages and warning supressed.
```{r}
library(caret)
library(kernlab)
library(randomForest)
library(rpart)
```

```{r, message=FALSE, warning=FALSE, cache=TRUE}
set.seed(1234)
inTrain <- createDataPartition(y=train3$classe, p=0.6, list=FALSE)
training <- train3[inTrain,]
testing <- train3[-inTrain,]

#rpart
mod.fit1 <- rpart(classe ~ ., data=training, method="class")
predict1 <- predict(mod.fit1, testing, type = "class")
confusionMatrix(predict1, testing$classe)

#randomForest
mod.fit2 <- randomForest(classe ~ ., data=training, method="class")
predict2 <- predict(mod.fit2, testing, type = "class")
confusionMatrix(predict2, testing$classe)
```


#####Conclusions
  RandomForest did a better job predicting the data so is used to test the independent data set.  The accuracy for rpart was 73% while the accuracy for randomForest was over 99.24%.  With this high of a percent for randomForest it is unlikely any test observeration will produce out ofsample errors.
  
  
#####Submission file creation
```{r}
predict.anwsers <- predict(mod.fit2, test3, type="class")

pml_write_files = function(x){
        n = length(x)
        for(i in 1:n){
                filename = paste0("problem_id_",i,".txt")
                write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
        }
}

        
pml_write_files(predict.anwsers)
```

The answers submitted are below.
```{r}
predict.anwsers
```


